# Learnr

Tutorials are available to run through with the R package [`learnr`](https://rstudio.github.io/learnr/).
These are interactive tutorials that give practice and exercises to supplement this site.

## Install instructions

You will need R and RStudio installed on your computer.

Within RStudio you will need to install the [`learnr`](https://rstudio.github.io/learnr/) package and the packages our [`learnr`](https://rstudio.github.io/learnr/) tutorials require.
Below are the R commands to install these.

```{r, eval=FALSE}
#learnr
install.packages("learnr")
#tidyverse
install.packages("tidyverse")
#lterdatasample
install.packages("lterdatasampler")
#Our packages are available through github rather than cran
#To install github based packages you need to install and load 
# the remotes package
install.packages("remotes")
library("remotes")
#mgrtibbles
remotes::install_github("m-gemmell/mgrtibbles")
#neoftidytutorials
#Package with the tidyverse learnr tutorials
remotes::install_github("m-gemmell/neoftidytutorials")
```

## Libraries

Prior to running any tutorials (next page) you must load the [`learnr`](https://rstudio.github.io/learnr/) and `neoftidytutorials` packages in your RStudio.

```{r, eval=FALSE}
library("learnr")
library("neoftidytutorials")
```