# Learnr tutorials

Tutorials are available to run through with the R package [`learnr`](https://rstudio.github.io/learnr/).
These are interactive tutorials that give practice and exercises to supplement this site.

## Install instructions

You will need R and RStudio installed on your computer.

Within RStudio you will need to install the [`learnr`](https://rstudio.github.io/learnr/) package and the packages our [`learnr`](https://rstudio.github.io/learnr/) tutorials require.
Below are the R commands to install these.

```{r, eval=FALSE}
#learnr
install.packages("learnr")
#tidyverse
install.packages("tidyverse")
#lterdatasample
install.packages("lterdatasampler")
#INstall hexbin
install.packages("hexbin")
#Our packages are available through github rather than cran
#To install github based packages you need to install and load 
# the remotes package
install.packages("remotes")
library("remotes")
#mgrtibbles
remotes::install_github("m-gemmell/mgrtibbles")
#neoftidytutorials
#Package with the tidyverse learnr tutorials
remotes::install_github("m-gemmell/neoftidytutorials")
```

## Libraries

Prior to running any tutorials (next page) you must load the [`learnr`](https://rstudio.github.io/learnr/) and `neoftidytutorials` packages in your RStudio.

```{r, eval=FALSE}
library("learnr")
library("neoftidytutorials")
```

## Tutorials

There are various tutorials to run through for different sections of this website.
Below shows you the order and how to run each tutorial.

Important notes about these tutorials:

- Progress is saved: If you close the tutorial and reopen it on the same computer & RStudio it will continue from where you left off
- Run in browser: You can click on the "Open in Browser" button at the top of the RStudio window with the tutorial to open the tutorial in a web browser
- Split screen: The intended way to run through the tutorials is with the tutorial on one half of your screen and this website at the relevant page on the other half

### Tibble

The first tutorial is on __Tibbles__.

To run the tutorial run the below [`learnr`](https://rstudio.github.io/learnr/) command.

__Note:__ Ensure you have followed the [install directions before this](learnr.qmd) and you have loaded the library (`library("neoftidytutorials")`).

```{r, eval=FALSE}
learnr::run_tutorial("tibbles", package = "neoftidytutorials")
```

Whilst running through the tutorial refer to the [Tibble section of this website](/tibble/tibble.qmd).

### Dplyr

The second tutorial is on the __`dplyr`__ package.

```{r, eval=FALSE}
learnr::run_tutorial("dplyr", package = "neoftidytutorials")
```

Whilst running through the tutorial refer to the [`dplyr` section of this website](/dplyr/dplyr.qmd).

### Tidyr

The third tutorial is on the __`tidyr`__ package.

```{r, eval=FALSE}
learnr::run_tutorial("tidyr", package = "neoftidytutorials")
```

Whilst running through the tutorial refer to the [`tidyr` section of this website](/tidyr/tidyr.qmd).

### Ggplot2

The fourth tutorial is on the __`ggplot2`__ package.

```{r, eval=FALSE}
learnr::run_tutorial("ggplot2", package = "neoftidytutorials")
```

Whilst running through the tutorial refer to the [`ggplot2` section of this website](/ggplot2/ggplot2.qmd).