# Tutorials

There are various tutorials to run through for different sections of this website.
Below shows you the order and how to run each tutorial.

Important notes about these tutorials:

- Progress is saved: If you close the tutorial and reopen it on the same computer & RStudio it will continue from where you left off
- Run in browser: You can click on the "Open in Browser" button at the top of the RStudio window with the tutorial to open the tutorial in a web browser
- Split screen: The intended way to run through the tutorials is with the tutorial on one half of your screen and this website at the relevant page on the other half

## Tibble

The first tutorial is on __Tibbles__.

To run the tutorial run the below [`learnr`](https://rstudio.github.io/learnr/) command.

__Note:__ Ensure you have followed the [install directions before this](learnr.qmd) and you have loaded the library (`library("neoftidytutorials")`).

```{r, eval=FALSE}
learnr::run_tutorial("tibbles", package = "neoftidytutorials")
```

Whilst running through the tutorial refer to the [Tibble section of this website](/tibble/tibble.qmd).

## Dplyr

The second tutorial is on the __`dplyr`__ package.

```{r, eval=FALSE}
learnr::run_tutorial("dplyr", package = "neoftidytutorials")
```

Whilst running through the tutorial refer to the [`dplyr` section of this website](/dplyr/dplyr.qmd).

## Tidyr

The third tutorial is on the __`tidyr`__ package.

```{r, eval=FALSE}
learnr::run_tutorial("tidyr", package = "neoftidytutorials")
```

Whilst running through the tutorial refer to the [`tidyr` section of this website](/tidyr/tidyr.qmd).