---
title: "Column types"
---

When reading in a file with `readr::delim()` the function will try to infer the type/class of data in each column.

More info on __tibbles__ and data classes in the [tibble structure page](/tibble/structure.qmd#variablescolumns).

Inferring the column types is not always correct and you can choose the column types with the `col_types=` option of the `reader::read_delim()` function.

Column types can be specified with a compact string of single character representations.
Each character represents the type of each column, with a number of strings equal to the number of columns (examples below).
The characters and their data type are:

- c = character
- i = integer
- n = number
- d = double
- l = logical
- f = factor
- D = date
- T = date time
- t = time
- ? = guess
- _ or - = skip

## Example
```{r, echo=FALSE}
#Vectors to be used as columns
Integers <- as.integer(c(101,21,3.3,0.04))
Doubles <- as.double(c(0.345,3.14,78.9,20000.9))
Characters <- c("One","sentence","is","enough")
Factors <- as.factor(c("A","B","B","A"))
Logicals <- c(TRUE,TRUE,FALSE,FALSE)
#Create a tibble and write to file as CSV
tibble::tibble(Integers, Doubles, Characters,Factors,Logicals) |>
    readr::write_delim("simple_example.csv", delim=",")
```

We'll read in a the file [simple_example.csv](https://neof-workshops.github.io/Tidyverse/data/simple_example.csv) to first specify column types.

### Check file contents

View the file contents before reading it as a __tibble__.

```{r}
readLines("https://neof-workshops.github.io/Tidyverse/data/simple_example.csv")
```

### Read in example file and specify column types

Read in the file specifying the column types.

```{r}
readr::read_delim(
    file = "https://neof-workshops.github.io/Tidyverse/data/simple_example.csv",
    delim = ",", 
    col_types = "idcfl")
```


