---
title: "Filter"
---

The `dplyr::filter()` function can be used to extract rows that meet a certain condition.

This can be a useful method to:

- Filter rows by numerics in a row, e.g. extract rows from years >= 2000.
- Extract rows belonging to certain groups of metadata.

[Tidyverse reference page](https://dplyr.tidyverse.org//reference/filter.html)

## Dataset

For demonstration we'll load the `knz_bison` data from the [lterdatasampler package](/datasets/lterdatasampler.qmd) (hyperlink includes install instructions).

```{r}
#Load package
library("lterdatasampler")
#knz_bison tibble for demonstration
bison_tbl <- lterdatasampler::knz_bison |>
    #Convert to tibble
    tibble::as_tibble()
bison_tbl
```

## Numeric columns

Filter the __tibble__ to only contain rows/observations from the year 2000 and onwards (>2000).

```{r}
bison_tbl |> dplyr::filter(rec_year > 2000)
```

## String columns

Filter the __tibble__ to only contain male observations.

```{r}
bison_tbl |> dplyr::filter(animal_sex == "M")
```

## Multiple columns

You can filter based on multiple columns with logical operators.

The 2 main R logical operators are:

- `&`: and (ampersand symbol)
- `|`: or (pipe symbol)

Extract females that have a weight greater or equal to 1000 pounds.

```{r}
bison_tbl |> dplyr::filter(animal_sex == "F" & animal_weight >= 1000)
```

Extract males or samples with a weight less than 900 pounds.

```{r}
bison_tbl |> dplyr::filter(animal_sex == "M" | animal_weight < 900)
```

## Pipes

You can use pipes for finer control of multiple comparisons.

The below command carries out the following filtering steps in order:

1. Extract females. 
2. Extract observations with animal weight between 900 and 1100. 
3. Extract samples from the year 2000.

```{r}
bison_tbl |> 
    #Filter to retain females
    dplyr::filter(animal_sex == "F") |>
    #Filter to retain samples with animal_weight between 900 and 1100
    dplyr::filter(animal_weight > 900 & animal_weight < 1000) |>
    #Retain samples from year 2000
    dplyr::filter(rec_year == 2000)
```