---
title: "NA removal"
---

You can remove rows with NAs in various ways with `dplyr`.

- Remove rows that contain at least one NA value
- Remove rows with NAs in one specific column
- Remove rows with NAs in certain columns

```{r, echo=FALSE, warning=FALSE,output=FALSE,error=FALSE}
library("tidyverse")
```

## Dataset

For demonstration we'll load the `mammal_sleep_tbl` data from the [mgrtibbles package](/datasets/mgrtibbles.qmd) (hyperlink includes install instructions).

```{r}
#Load package
library("mgrtibbles")
#mammal_sleep_tbl tibble for demonstration
mgrtibbles::mammal_sleep_tbl
```

## Na in any column

Remove rows with an NA in any column with `na.omit()` from `stats` package in base R.

```{r}
mammal_sleep_tbl |>
    na.omit()
```

## NA in one column

Can remove rows with an NA in one specific column with `dplyr::filter()` and `!is.na()`.

Remove rows where there is an NA in the `life_span` column.

```{r}
mammal_sleep_tbl |>
    dplyr::filter(!is.na(life_span))
```

## NA in certain columns

Remove rows with an NA in at least one of the specified columns with `dplyr::filter_at()`.

Remove rows where there is an NA in the non_dreaming or dreaming columns.

```{r}
mammal_sleep_tbl |>
    dplyr::filter_at(vars(non_dreaming, dreaming), all_vars(!is.na(.)))
```

