---
title: "Rename"
---

The function `dplyr::select()` allows you to select columns from a __tibble__.
There are many different ways to do this with various helper functions.

When subsetting with `dplyr::select()` the resulting object will always be a __tibble__.

[Tidyverse reference page](https://dplyr.tidyverse.org//reference/select.html)

## Dataset

For demonstration we'll load the `mammal_sleep_tbl` data from the [mgrtibbles package](/datasets/mgrtibbles.qmd) (hyperlink includes install instructions). For easier viewing we'll subset it so it only has 5 rows.

```{r}
#Load package
library("mgrtibbles")
#mammal_sleep_tbl tibble for demonstration
mammal_sleep_tbl<- mgrtibbles::mammal_sleep_tbl |> dplyr::slice(1:5)
mammal_sleep_tbl
```

# Rename columns

The format of the `rename()` function is:

```{r, eval=FALSE}
dplyr(new_name=old_name)
```

Rename column __total_sleep__ to __tot_sleep__. 

```{r}
mammal_sleep_tbl |> dplyr::rename(tot_sleep=total_sleep)
```

Rename columns __body_wt__ and __brain_wt__ to __body_kg__ and __brain_kg__.

```{r}
mammal_sleep_tbl |> dplyr::rename(body_kg=body_wt, brain_kg=brain_wt)
```