---
title: "Printing or viewing"
format: html
---

Tibbles have some unique characteristics when printing and viewing them on a terminal, including the RStudio console window.

## Load and print data

Below we load the `and_vertebrates` data from the [lterdatasampler package](/datasets/lterdatasampler.qmd) (includes install instructions) and view it.

```{r}
#Load package
library("lterdatasampler")
#subsetted pie_crab tibble
tibble::as_tibble(and_vertebrates)[1:10,1:5]
```

Information on the format of a __tibble__ is within the [Tibble structure section](/tibble/structure.html#format)

## Larger tibbles

When printing out a larger __tibble__ only a part of the __tibble__ will be displayed with info on the undisplayed data below the displayed data.

By default only 10 __rows__ of a __tibble__ will be displayed. 

The number of displayed __columns__ is dynamic.
The maximum amount of __columns__ will be displayed based on the width of the viewer and the width of the data in the __columns__.

The info shown below a __tibble__ is:

- The number of rows that are not displayed
- The number, names, and classes of the variables/columns not displayed

```{r}
tibble::as_tibble(and_vertebrates)
```

## Print more rows

More rows can be printed by using the base R `print()` function with the `n=` option.

```{r}
print(tibble::as_tibble(and_vertebrates), n =20)
```

