---
title: "Subsetting"
format: html
---

There are various ways to subset a __tibble__. This sections shows the various methods with examples.

[Tidyverse's tibble subsetting webpage](https://tibble.tidyverse.org/reference/subsetting.html)

## Load data

The examples in this section use the `pie_crab` data from the [lterdatasampler package](/datasets/lterdatasampler.qmd) (includes install instructions).

Load package
```{r}
library("lterdatasampler")
```

Create a __tibble__ from a subset of the pie_crab data.
```{r}
pie_crab_tbl <- tibble::as_tibble(pie_crab[1:10,1:8])
pie_crab_tbl
```

## One column with $

Subsetting a column using `$` and a column name will produce a __vector__.

```{r}
pie_crab_tbl$site
str(pie_crab_tbl$site)
```

## Columns with [c]

Subsetting 1 or more columns with `[c]` will produce a __tibble__.

```{r}
pie_crab_tbl[1]
```

```{r}
pie_crab_tbl[c("date","air_temp","size")]
```

## Rows and columns with [r,c]

Subsetting a __tibble__ with `[]` will always produce a __tibble__.

This is different from a __data.frame__ where you may extract a [scalar](https://neof-workshops.github.io/R_j4c0xh/04-R_objects.html#scalar), [vector](https://neof-workshops.github.io/R_j4c0xh/04-R_objects.html#vectors-lists), or __data.frame__.

Below are examples using a subset of the `pie_crab` data from the [lterdatasampler package](/datasets/lterdatasampler.qmd) (includes install instructions).

### Load data.frame

Create a __data.frame__ from the `pie_crab_tbl`.
```{r}
pie_crab_df <- as.data.frame(pie_crab_tbl)
pie_crab_df
```

### Extracting a value

Subsetting one value from a __data.frame__ leads to a __scalar__
```{r}
pie_crab_df[1,1]
str(pie_crab_df[1,1])
```

Subsetting one value from a __tibble__ leads to a __tibble__
```{r}
pie_crab_tbl[1,1]
```

### Extracting a row

Subsetting a row from a __data.frame__ leads to a __data.frame__

```{r}
pie_crab_df[1,]
str(pie_crab_df[1,])
```

Subsetting a row from a __tibble__ leads to a __tibble__
```{r}
pie_crab_tbl[1,]
```

### Extracting a column

Subsetting a column from a __data.frame__ leads to a __vector__
```{r}
pie_crab_df[,1]
str(pie_crab_df[,1])
```

Subsetting a column from a __tibble__ leads to a __tibble__
```{r}
pie_crab_tbl[,1]
```

### Extracting a combo of columns and rows

Subsetting rows and columns together from a __data.frame__ leads to a __data.frame__
```{r}
pie_crab_df[1:3,c("date","air_temp","size")]
str(pie_crab_df[1:3,c("date","air_temp","size")])
```

Subsetting rows and columns together from a __tibble__ leads to a __tibble__
```{r}
pie_crab_tbl[1:3,c("date","air_temp","size")]
```
