---
title: "Tibble structure"
format: html
---

Useful resource for MG: https://ssc.wisc.edu/sscc/pubs/R_intro/book/2-5-tibble-structure.html

Tibbles are formatted like a __data.frame__ but have some extra functions and terminology.

## Format

Below is an example of a tibble (Dataset: [Sleep in mammals](/datasets/mammal_sleep.qmd)).

```{r, eval=FALSE}
mammal_sleep <- readr::read_csv(file = "mammals.csv", show_col_types = FALSE)
mammal_sleep[1:10,1:6]
```

```{r, echo=FALSE}
mammal_sleep <- readr::read_csv(file = "../datasets/mammals.csv", show_col_types = FALSE)
mammal_sleep[1:10,1:6]
```

Let's go over some of its components:

- Size info
- Variables/columns
- observations/rows

### Size info

At the top of the __tibble__ it tells us the size.
The first number is the rows and the second is the columns.

In the above example we have:

- 10 rows/observations
- 6 columns/variables

### Rows/observations

As in a __data.frame__ each observation is within its own row.
The contents of an observation/row is determined by your dataset and how you want to analyse or visualise it.
There are many ways to manipulate __tibble__ with the [dplyr](/dplyr/dplyr.qmd), [tidyr](/tidyr/tidyr.qmd), and other tidyverse packages. 

__Tibble__ rows do not have row names and you cannot/should not give them row names.
They are numbered instead.

### Variables/columns

A __tibble__ will have a certain number of columns.

Each column has a name that is at the top. 
For example:

- species
- body_wt
- brain_wt

Below the column/variable names is the data class/type of the column.
All the entries within a column will have the same data class.
These include:

- __\<int>__: [Integer](https://neof-workshops.github.io/R_j4c0xh/04-R_objects.html#numeric)
- __\<dbl>__: [Double](https://neof-workshops.github.io/R_j4c0xh/04-R_objects.html#numeric)
- __\<chr>__: [Character/String](https://neof-workshops.github.io/R_j4c0xh/04-R_objects.html#string)
- __\<fct>__: [Factor](https://neof-workshops.github.io/R_j4c0xh/04-R_objects.html#string)
- __\<lgl>__: [Logical](https://neof-workshops.github.io/R_j4c0xh/04-R_objects.html#string)
- __\<date>__: Date (MG ADD LINK TO INFO ABOUT DATE CLASS, MOST LIKELY IN LUBRIDATE SECTION)
- __\<list>__: List

Below is an example of a __tibble__ with column for each of these classes.

```{r, echo = FALSE}
#Vectors to be used as columns
Integers <- as.integer(c(101,21,3.3,0.04))
Doubles <- as.double(c(0.345,3.14,78.9,20000.9))
Characters <- c("One","sentence","is","enough")
Factors <- as.factor(c("A","B","B","A"))
Logicals <- c(TRUE,TRUE,FALSE,FALSE)
Dates <- as.Date(c("01-01-2020", "02-01-2020", "03-01-2020","04-01-2020"))
Lists <- list(c(1,2,3),c(2,3,4),c(3,4,5),c(4,5,6))

#Create a tibble
tbl <- tibble::tibble(Integers, Doubles, Characters,Factors,Logicals,Dates, Lists)
#View tibble
#Note that the column names are the same as the vector names
tbl
```

Above you can see that when you view a __tibble__, by printing it or viewing it in an Rstudio tab, it will align the values in columns in a easy to read manner.
This is especially useful for numeric (integers & doubles) columns where the numbers are aligned by place value.
This allows you to easily compare the size of numbers in the same column.